Robust Control of a Fluxonium Qubit

### 2020-11-28 paper/1introduction.tex
solid progress through sec4

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
\todo{more nmr and sqc cite} Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{
  brouzos2015quantum, de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
  treutlein2006microwave, van2016optimal}, nitrogen-vacancy centers in
diamond \cite{chou2015optimal, dolde2014high, geng2016experimental,
  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
For quantum computation,
optimal control techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system to deviate
from the model used in optimization, leading
to poor experimental performance.
Robust control techniques improve upon
standard optimal control techniques by encoding
model uncertainties
in optimization objectives, yielding performance
guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
We adapt robust control techniques from the robotics community to mitigate
realistic model uncertainties for
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have had tremendous success in
engineering high fidelity quantum gates \cite{
  chakram2020multimode, chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
  gokhale2019partial, grace2007optimal, heeres2017implementing, huang2014optimal,
  kelly2014optimal, leng2019robust, liebermann2016optimal,
  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
  spiteri2018quantum, sporl2007optimal}. 
While traditional QOC techniques can predict system behavior
with high accuracy, they are sensitive to experimental
errors such as parameter drift, noise, finite control 
resolution, and decoherence. Analytic techniques to mitigate parameter deviation errors
include dynamic and geometric phase considerations
\cite{han2020experimental, xu2020nonadiabatic} and
composite pulses \cite{merrill2014progress}.
To mitigate decoherence, Floquet techniques have been employed
\cite{huang2020engineering, mundada2020floquet}.
Numerical techniques to mitigate parameter deviation errors
include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
  wittler2020integrated} and open-loop methods \cite{
  allen2019robust, carvalho2020error, reinhold2019controlling,
  rembold2020introduction, kosut2013robust}.
Numerical techniques to mitigate decoherence include
modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.

>>>>

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{vandersypen2005nmr, kehlet2004improving, khaneja2005optimal,
  maximov2008optimal, nielsen2010optimal, skinner2003application, tosner2009optimal},
and applications now include superconducting circuits \cite{abdelhafez2020universal,
  chakram2020multimode, fisher2010optimal, gokhale2019partial,
  huang2020engineering, leng2019robust, leung2017speedup, li2020fast,
  xu2020nonadiabatic},
neutral atoms and ions \cite{brouzos2015quantum,
  de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
  treutlein2006microwave, van2016optimal},
nitrogen-vacancy centers in diamond \cite{chou2015optimal,
  dolde2014high, geng2016experimental,
  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
For quantum computation,
optimal control techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system to deviate
from the model used in optimization, leading
to poor experimental performance.
Robust control techniques improve upon
standard optimal control techniques by encoding
model uncertainties
in optimization objectives, yielding performance
guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
We adapt robust control techniques from the robotics community to mitigate
realistic model uncertainties for
a superconducting fluxonium qubit.

QOC techniques have had tremendous success in
engineering high fidelity quantum gates \cite{
  chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
  grace2007optimal, heeres2017implementing, huang2014optimal,
  kelly2014optimal, leng2019robust, liebermann2016optimal,
  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
  spiteri2018quantum, sporl2007optimal}. 
While standard QOC techniques can predict system behavior
with high accuracy, they are sensitive to experimental
errors such as parameter drift, noise, finite control 
resolution, and decoherence. Multiple techniques have been developed to address
these shortcomings. Analytic techniques to mitigate parameter deviation errors
include dynamic and geometric phase considerations
\cite{han2020experimental, xu2020nonadiabatic} and
composite pulses \cite{cummins2000use, cummins2003tackling,
  kupce1995stretched, merrill2014progress}.
To mitigate decoherence, Floquet techniques have been employed
\cite{huang2020engineering, mundada2020floquet}.
Numerical techniques to mitigate parameter deviation errors
include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
  wittler2020integrated} and open-loop methods \cite{
  allen2019robust, carvalho2020error, reinhold2019controlling,
  rembold2020introduction, kosut2013robust}.
Numerical techniques to mitigate decoherence include
modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.

### 2020-11-28 paper/2background.tex
solid progress through sec4

\section{QOC + AL-iLQR \label{sec:background}}
In this section we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
  \label{eq:tdse}
\end{equation}
The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
and Runge-Kutta integrators \cite{jorgensen2011numerical}.

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u(t))} \rvert}^{2}$.
In general $J$ is a linear combinaion of cost functions on the state
as well as cost functions on the controls \cite{leung2017speedup}.
Typical quantum optimal control
algorithms employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{clarkson2010coresets, hauswirth2016projected}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
$u_k \in \R^m$ are the state and input control variables, $f(x_k,u_k)$ is the discrete
dynamics function, and $g_k(x_k,u_k)$ and $h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

@@

The ALTRO method has fast tail-convergence and
allows us to impose arbitrary constraints without restricting
the optimization to the constraint manifold.
This contrasts typical first-order QOC methods that employ the same
update procedure for the duration of optimization
and employ projected gradient methods
which restrict the optimization to the constraint manifold.
To see how to use the ALTRO method for QOC, we will present
a concrete example in the following section.

>>>>

\section{QOC + ALTRO \label{sec:background}}
In this section we
review the quantum optimal control problem statement
and introduce the ALTRO solver.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}
  \label{eq:tdse}
\end{equation}
The Hamiltonian has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. To make the problem numerically tractable,
the controls and quantum state are discretized into $N$ knot points (time steps).
The optimization problem is to find the
controls that minimize a functional $J(u)$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
In general, $J$ is a linear combinaion of cost functions on the state
as well as cost functions on the controls \cite{leung2017speedup}.
Typical quantum optimal control
algorithms employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{machnes2015gradient, khaneja2005optimal, leung2017speedup, goerz2019krotov}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
$u_k \in \R^m$ are the augmented state and input augmented control variables,
$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
$h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

@@

As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
for imposing arbitrary, simultaneous constraints
on both the state trajectory and the controls and does not restrict
optimization to the constraint manifold. First-order solvers like GRAPE
rely on unconstrained optimization to satisfy desiderata on the state trajectory,
such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
and smoothing the controls below a threshold. The unconstrained approach relies
on a careful choice of competing penalty weights, which is cumbersome and often intractable.
To impose constraints on the controls, such as amplitude restrictions, first-order solvers
employ projected gradient descent methods which restrict optimization to the constraint manifold and
may hinder convergence to the optimal solution \cite{clarkson2010coresets,
  hauswirth2016projected}.
The ALTRO solver delays projecting onto the constraint manifold until the ALM
solve is already at a coarse tolerance, achieving machine-precision satisfaction
for constraints on the controls and the state trajectory. These advantage are critical
for this work, where multiple constraints must be satisfied in addition
to minimizing auxiliary metrics.

### 2020-11-28 paper/3fluxonium.tex
solid progress through sec4

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
  {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{(x_1, x_N \; \textrm{given})}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
\end{mini!}
where the augmented state and controls are given by:
\begin{equation}
  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The initial and final conditions are determined by the
desired unitary transformation, which requires two quantum states
to specify $\psi^{(1)}$, $\psi^{(2)}$.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
given in \cite{leung2017speedup}:
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
\end{equation}
The discrete dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
\eqref{eq:tdse}, \eqref{eq:hamiltonian}
and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric. We choose $Q$
to be diagonal because it is computationally efficient. This corresponds
to penalizing phase differences between the quantum states, although
other phase-insensitive metrics such as infidelity may be employed.
The final state transfer is the most important objective which is
encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\psi_{N} = \psi_{f}$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$.

%% In addition to obeying experimental constraints
%% and achieving low simulated gate errors, we desire
%% to make our gates perform well in the presence of decoherence.
%% Decoherence of the quantum state due to external noise
%% is modeled by two phenomena: longitudinal relaxation and pure dephasing.
%% They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
%% Dissipation to the thermal bath via longitudinal
%% relaxation is an irreversible process
%% that results in information loss.
%% Converesely, pure dephasing is a reversible process.
%% We choose to maximize the longitudinal relaxation time directly
%% by tuning the flux amplitude, as in Section \ref{sec:longitude}, and
%% employ robust control techniques to mitigate
%% pure dephasing in Section \ref{sec:stochastic}.

%% The main contributions to longitudinal relaxation in our
%% device are dielectric loss in the capacitor, resistive loss in the inductor,
%% and Purcell loss. The main contributions to pure dephasing in our
%% device are $1/f$ flux noise and decay via charge and flux coupling
%% to the control lines.
%% Our device achieves its best pure dephasing
%% protection at the flux frustration point
%% $T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
%% where the qubit is first-order insensitive to changes in flux.
%% It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
%% frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
%% at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
%% and increases away from the flux frustration point
%% $T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
%% of the decay processes and the tradeoff,

>>>>

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the accurate
two-level approximation of its Hamiltonian makes
quantum optimal control on a classical computer inexpensive.
In the two-level
approximation the Hamiltonian takes the form:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a(t)$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
Hamiltonian parameter deviations. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
  {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{(x_1, x_N \; \textrm{given})}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
\end{mini!}
where the augmented state and controls are given by:
\begin{equation}
  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The initial and final conditions are determined by the
desired gate, which requires two quantum states
to specify $\psi^{(1)}$, $\psi^{(2)}$.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
given in \cite{leung2017speedup}:
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
  \label{eq:isomorphism}
\end{equation}
The discrete dynamics function \eqref{eq:dyn_con} integrates the TDSE dynamics for the quantum states
using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
Including objectives at each knot point smoothens the optimization landscape, though
the importance of the final state objective is encoded in the relative
weight $Q_{N} \sim N \cdot Q_{k}$.
We choose $Q$ to be diagonal because it is computationally efficient. This corresponds
to penalizing phase differences between the quantum states, although
other phase-insensitive metrics such as infidelity may be employed.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\psi_{N} = \psi_{f}$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition, the norm of the quantum states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
of subsets from $x_{k}$ is implied in
the constraint equations.

### 2020-11-28 paper/4longitude.tex
solid progress through sec4

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
    \caption{\label{fig:longitudea}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
    \caption{\label{fig:longitudeb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
    \caption{\label{fig:longitudec}}
  \end{subfigure}
  \caption{
    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The cumulative
    gate error is computed after each gate application.
    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
    in the figure.
  }
  \label{fig:longitude}
\end{figure*}

\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
Dissipation to the thermal bath via longitudinal relaxation
is an irreversible process that results in information loss.
This process is modeled using its $1/e$ decay time $T_{1}$.
The longitudinal relaxation time $T_{1}$ varies with
control parameters in a range of superconducting circuit platforms,
including the fluxonium qubit.
It is advantageous to tune the controls to extend the longitudinal
relaxation time.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics \eqref{eq:tdse}.
We avoid this increase in computational complexity by
penalizing the integrated rate of longitudinal relaxation,
i.e. the probability of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation
increase monotonically in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

The longitudinal relaxation probability is given by,
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
This value is appended to the augmented state vector \eqref{eq:astatecontrols}
and penalized according to \eqref{eq:costfun}.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.
It is also possible to use a spline fit to theoretically obtained data.
However, $T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data
increases the fridge truth of the simulation.

@@

We compare the numerical method we have developed to the analytic gates
on the task of achieving low gate errors in the presence of longitudinal relaxation
for the $Z/2$, $Y/2$, and $X/2$ gates.
The numerically optimized gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitude}.
They extend their gate times
beyond their analytic counterparts, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
their single gate errors by a factor of $5$ over
their analytic counterparts which is commensurate to their
probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The numerical $X/2$ gate
achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
Both the analytic and numerical gates attain single gate errors sufficient for
quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
The low gate errors achieved by the numerical gates \todo{typo / awkward sentence} 
for extend computations
are critical for noisy, intermediate-scale quantum (NISQ) applications.
These improvements are significant for the realistic constraints we have imposed
on the gates, and do not represent a fundamental limit to the optimization methods we have
employed.

>>>>

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
    \caption{\label{fig:longitudea}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
    \caption{\label{fig:longitudeb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
    \caption{\label{fig:longitudec}}
  \end{subfigure}
  \caption{
    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The cumulative
    gate error is computed after each gate application.
    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable.
  }
  \label{fig:longitude}
\end{figure*}

\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
The strength of longitudinal relaxation type decoherence varies
with the control parameters in a range of superconducting circuit platforms,
including the fluxonium qubit. Because dissipation to the thermal bath
via longitudinal relaxation is an irreversible process that results in information loss,
it is advantageous to tune the controls to extend the longitudinal
relaxation time $T_{1}$--which represents the $1/e$ decay time
of the quantum state.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics \eqref{eq:tdse}.
We avoid this increase in computational complexity by
penalizing the probability (integrated rate) of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation increase monotonically
in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

The longitudinal relaxation probability is given by,
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
This value is appended to the augmented state vector \eqref{eq:astatecontrols}
and penalized according to \eqref{eq:costfun}.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.
It is also possible to fit a spline to theoretically obtained data.
However, $T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data
increases the fridge truth of the simulation.

@@

We compare the numerical method we have developed to the analytic gates
on the task of achieving low gate errors in the presence of longitudinal relaxation
for the $Z/2$, $Y/2$, and $X/2$ gates.
The numerically optimized gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They extend their gate times
beyond their analytic counterparts, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
their single gate errors by a factor of $5$ over
their analytic counterparts which is commensurate to their
probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
the concatenated gate application comparision, suppressing cumulative gate errors to $8 \cdot 10^{-3}$
over $2000$ gate applications $\sim 40 \mu\textrm{s}$, see Figure \ref{fig:longitudec}.
The numerical $X/2$ gate achieves a cumulative gate error of
$1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
These low cumulative gate errors for high gate counts are critical for
noisy, intermediate-scale quantum (NISQ) applications.
Both the analytic and numerical gates attain single gate errors sufficient for
quantum error correction ($< 10^{-4}$), which are required for fault-tolerant quantum computing
\cite{aharonov2008fault, fowler2009high, gottesman1997stabilizer}.
These improvements are significant for the realistic constraints we have imposed
on the gates, and do not represent a fundamental limit to the optimization methods we have
employed.
